Data preparation

### 2020-09-10 docs/index.Rmd

---
title: "PPM-S"
csl: "../input/bib/apa6.csl"
bibliography: "../input/bib/zlibrary.bib"
author: " "
output:
    html_document:
        smart: false
        includes:
          in_header: head_github-activity.html
---

>>>>

---
title: "PPM-S"
csl: "../input/bib/apa6.csl"
bibliography: "../input/bib/zlibrary.bib"
author: " "
output:
    bookdown::html_document2:
            smart: false
            includes:
              in_header: head_github-activity.html
            number_sections: false          
---

### 2020-09-10 docs/paper.Rmd

```{r eval=FALSE, include=FALSE}
# for render in pdf run rmarkdown::render_site("docs/paper.Rmd", output_format = "all")
# clean #in the yml
rmarkdown::render("docs/paper.Rmd", output_format = "bookdown::pdf_document2")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = FALSE,out.width = '75%',fig.pos= "H")
# knitr::opts_knit$set(base.url = "../") #relative path for .html output file
# knitr::opts_knit$set(root.dir = "../") #relative path for chunks within .rmd files
options(scipen=999)
rm(list=ls())
options(knitr.kable.NA = '')
options(knitr.graphics.error = FALSE)
Sys.setlocale("LC_ALL", "ES_ES.UTF-8")
```

@@

kable(tab01,table_format,booktabs = T, linesep = "",col.names = cnames, caption = cap) %>%
  kable_styling(
    full_width = T,
    latex_options = c("hold_position", "scale_down"),
    position = "center",
    font_size = 10,
    bootstrap_options=c("striped", "bordered")) %>%
  column_spec(column = 1, width = "1.5cm", ) %>%
  column_spec(column = 2,width = "2 cm") %>%
  collapse_rows(columns = 1:2,valign = "middle")
```

@@

```{r desc01, echo=FALSE}
cap <- "Descriptive statistics of the scale."
load(file = "../input/data/proc/dat04.RData")
dat04b<- dat04 %>% rename("A. Perception Effort" =perc_effort,
                          "B. Perception Talent" = perc_talent,
                          "C. Perception rich parents" = perc_wpart,
                          "D. Perception contacts" = perc_netw,
                          "E. Preferences Effort" = pref_effort,
                          "F. Preferences Talent" =  pref_talent,
                          "G. Preferences rich parents" =  pref_wpart ,
                          "H. Preferences contacts" = pref_netw  )
sum04b  <- psych::describe(dat04b[,-c(1)], fast = T) %>% select(mean,sd,min,max) %>% data.frame() %>% round(digits = 2)
table04b<- sum04b %>% rename("Mean"=mean, "SD"=sd, "Min"=min,"Max"=max)
sumdat04<- kable(table04b,format =table_format, booktabs= T, escape = FALSE,align = "l",linesep = "", caption = cap
) %>% kable_styling(full_width = table_format2, font_size = 10,position = "center",latex_options = c("hold_position"));sumdat04
```

@@

```{r meas02, echo=FALSE, fig.cap = "Theorical model", fig.align='center'}
knitr::include_graphics('../output/images/meas02.png')
```

>>>>

```{r eval=FALSE, include=FALSE}
# for render in pdf run rmarkdown::render_site("docs/paper.Rmd", output_format = "all")
# clean #in the yml
rmarkdown::render("docs/paper.Rmd", output_format = "bookdown::pdf_document2")
rmarkdown::render("docs/paper.Rmd", output_format = "bookdown::html_document2")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = FALSE,out.width = '85%',fig.pos= "H")
# knitr::opts_knit$set(base.url = "../") #relative path for .html output file
# knitr::opts_knit$set(root.dir = "../") #relative path for chunks within .rmd files
options(scipen=999)
rm(list=ls())
options(knitr.kable.NA = '')
options(knitr.graphics.error = FALSE)
Sys.setlocale("LC_ALL", "ES_ES.UTF-8")
```

@@

kable(tab01,table_format,booktabs = T, linesep = "",col.names = cnames, caption = cap) %>%
  kable_styling(
    full_width = T,
    latex_options = c("hold_position"),
    position = "center",
    font_size = 10,
    bootstrap_options=c("striped", "bordered")) %>%
  column_spec(column = 1, width = "1.5cm", ) %>%
  column_spec(column = 2,width = "2 cm") %>%
  collapse_rows(columns = 1:2,valign = "middle")
```

@@

```{r desc01, echo=FALSE}
cap <- "Descriptive statistics of the scale."
load(file = "../input/data/proc/dat04.RData")
dat04b<- dat04 %>% rename("A. Perception Effort" =perc_effort,
                          "B. Perception Talent" = perc_talent,
                          "C. Perception rich parents" = perc_wpart,
                          "D. Perception contacts" = perc_netw,
                          "E. Preferences Effort" = pref_effort,
                          "F. Preferences Talent" =  pref_talent,
                          "G. Preferences rich parents" =  pref_wpart ,
                          "H. Preferences contacts" = pref_netw  )
sum04b  <- psych::describe(dat04b[,-c(1)], fast = T) %>% select(mean,sd,min,max) %>% data.frame() %>% round(digits = 2)
table04b<- sum04b %>% rename("Mean"=mean, "SD"=sd, "Min"=min,"Max"=max)
sumdat04<- kable(table04b,format =table_format, booktabs= T, escape = FALSE,align = "l",linesep = "", caption = cap
) %>% kable_styling(full_width = table_format2, 
                    font_size = 10,
                    position = "center", 
                    latex_options = c("hold_position"));sumdat04
```

@@

```{r meas02, echo=FALSE, fig.cap = "Confirmatory factor analysis of the Perceptions an Preferences for Meritocracy Scale", fig.align='center'}
knitr::include_graphics('../output/images/meas02.png')
```

### 2020-09-10 production/prod_analysis-cfa.rmd

**Libraries**
```{r}
library(sjPlot)
library(dplyr)
library(lavaan)
library(semPlot)
library(stargazer)
library(corrplot)
library(psych)
library(knitr)
library(kableExtra)
library(rvest)
library(sjlabelled)
source('production/prod_polychoric.R')
```

@@

```{r}
cor04<-polychoric(select(dat04,perc_effort,perc_talent,perc_wpart,perc_netw,pref_effort,pref_talent,pref_wpart,pref_netw))
rownames(cor04$rho) <-c("A.Perception Effort", "B.Perception Talent","C.Perception Rich parents","D.Perception contacts","E.Preferences Effort","F.Preferences Talent","G.Preferences rich parents", "H.Preferences contacts" )
colnames(cor04$rho) <-c("(A)", "(B)","(C)","(D)","(E)","(F)","(G)", "(H)")
save(cor04, file = "output/images/correlation.RData")
corrplot(cor04$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = TRUE)

png("output/images/corpoly.png",
    width = 15,height = 15,units="cm",
    pointsize=12,bg="white",res=300)

corrplot(cor01$rho,method = "color" ,type = "upper", tl.col = "black", addCoef.col = "black",diag = TRUE)

@@

semPaths(
  fit4_o ,
  whatLabels = "std",
  # what = "col",
  label.cex = 1,
  edge.label.cex = 0.7,
  residuals = F,
  optimizeLatRes =T,
  edge.color = "black",
  style = "lisrel",
  fade = F,
  nCharNodes = 0,
  curvePivot = FALSE,
  curve = 3,
  rotation = 4,
  layout = "tree2",
  cardinal = "lat cov",
  legend.cex = 0.6,
  label.cex = 1,
  label.font = 6,
  edge.label.font = 7,
  asize = 3,
  edge.width = 1,
  sizeMan = 5,
  sizeLat = 10,
  width = 20,
  height = 15,
  nodeNames = nodeNames,
  intercepts = F,
  reorder = F,
  thresholds = F,
  fixedStyle =1,
  nodeLabels = c("A", "B", "C", "D", "E", "F", "G", "H", "1", "2", "3", "4"),
  filetype = "png",
  filename = "output/images/meas02"
) #arrow size)

@@

>>>>

**Libraries**
```{r}
library(sjPlot)
library(dplyr)
library(lavaan)
library(semPlot)
library(stargazer)
library(corrplot)
library(psych)
library(knitr)
library(kableExtra)
library(rvest)
library(sjlabelled)
library(ggplot2)
source('production/prod_polychoric.R')
```

@@

```{r}
cor04<-polychoric(select(dat04,perc_effort,perc_talent,perc_wpart,perc_netw,pref_effort,pref_talent,pref_wpart,pref_netw))
windowsFonts(A = windowsFont("Times New Roman"))
rownames(cor04$rho) <-c(
    "A. Perception Effort",
    "B. Perception Talent",
    "C. Perception Rich parents",
    "D. Perception Contacts",
    "E. Preferences Effort",
    "F. Preferences Talent",
    "G. Preferences Rich parents",
    "H. Preferences Contacts")
colnames(cor04$rho) <-c("(A)", "(B)","(C)","(D)","(E)","(F)","(G)", "(H)")
save(cor04, file = "output/images/correlation.RData")
corrplot(
  cor04$rho,
  method = "color",
  type = "upper",
  tl.col = "black",
  addCoef.col = "black",
  diag = TRUE,
  family = "A", 
  number.font = 6,
  tl.cex =0.75,
  number.cex = 1)

png("output/images/corpoly.png",
    width = 18,height = 13,units="cm",
    pointsize=12,bg="white",res=300)

corrplot(
  cor04$rho,
  method = "color",
  type = "upper",
  tl.col = "black",
  addCoef.col = "black",
  diag = TRUE,
  family = "A", 
  number.font = 6,
  tl.cex =0.75,
  number.cex = 1)

@@

semPaths(
  fit4_o ,
  whatLabels = "std",
  # what = "col",
  label.cex = 1,
  edge.label.cex = 0.7,
  residuals = F,
  optimizeLatRes =T,
  edge.color = "black",
  style = "lisrel",
  fade = F,
  nCharNodes = 0,
  curvePivot = FALSE,
  curve = 3,
  rotation = 4,
  layout = "tree2",
  cardinal = "lat cov",
  legend.cex = 0.6,
  label.cex = 1,
  label.font = 6,
  edge.label.font = 7,
  asize = 3,
  edge.width = 1,
  sizeMan = 5,
  sizeLat = 10,
  width = 22, 
  height = 15,
  # nodeNames = nodeNames,
  intercepts = F, 
  reorder = F,
  thresholds = F,
  fixedStyle =1,
  nodeLabels = c("A", "B", "C", "D", "E", "F", "G", "H", "1", "2", "3", "4"),
  filetype = "png",
  filename = "output/images/meas02"
) #arrow size)
# text(2.5, -1, labels = "Hola")
```

@@

```{r}
datli03 <- dat04 %>% select(starts_with("perc_"),
                            starts_with("pref_")) %>%  sjmisc::rec(rec="rev",append = F)

datli03 <- remove_label(datli03)

set_theme(base = theme_bw(),
          theme.font = "serif", 
          axis.textsize.x = 1.2,
          axis.textsize.y = 1.2, 
          geom.label.size = 4, 
          legend.just = 0.35, 
          legend.size = 1.2,
          legend.backgroundcol = "white")
# http://www.sthda.com/english/wiki/colors-in-r  = PALETAS de COLORES 
plot01<- plot_likert(
  datli03,
  c(rep("Perception", 4), rep("Preference", 4)),
  values  =  "sum.outside",
  show.prc.sign = F,
  sort.groups = FALSE,
  grid.range = c(1.0, 1.0),
  geom.colors = "Blues",
  rel_heights = c(4, 4),
  wrap.labels = 100,
  reverse.scale = FALSE,
  cat.neutral = 3, 
  legend.pos = "bottom", 
  group.legend.options = list(reverse = TRUE),
  axis.labels = c("Effort", "Talent", "Rich parents", "Contacts"))
plot01

ggsave(plot01,filename = "output/images/plotlikert.png",device = "png",width = 40,height = 25,dpi = "retina",units = "cm")
```

### 2020-09-10 production/prod_prep-data.rmd

## Translation of variable labels and value labels.
```{r}
label(dat04$grupo) <- "This variable classifies the interviewees according to the order in which the indicators are shown within the scale. There are three orders  of items on the scale. In the first order (group 1), questions of the same factor are then presented to them, in the second order (group 2) they were separated and in the third order (group 3) at random."
dat04$perc_effort <- set_labels(dat04$perc_effort,  labels = c("strongly disagree" = 1, "disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "totally agree" = 5))
label(dat04$perc_effort) <- "Thinking of Chilean society,
what measure do you find of
agree or disagree with each of
the following statements? - Who
the more they try they manage to get bigger
rewards that those who strive
less."
dat04$perc_talent <- set_labels(dat04$perc_talent,  labels = c("strongly disagree" = 1, "disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "totally agree" = 5))
label(dat04$perc_talent) <- "Thinking of Chilean society,
what measure do you find of
agree or disagree with each of
the following statements? - Who
possess more talent they manage to obtain
greater rewards than those who possess
less talent."
dat04$perc_wpart <- set_labels(dat04$perc_wpart,  labels = c("strongly disagree" = 1, "disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "totally agree" = 5))
label(dat04$perc_wpart) <- "Thinking of Chilean society,
what measure do you find of
agree or disagree with each of
the following statements? - Who
they have rich parents manage to get out
ahead."
dat04$perc_netw <- set_labels(dat04$perc_netw,  labels = c("strongly disagree" = 1, "disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "totally agree" = 5))
label(dat04$perc_netw) <- "Thinking of Chilean society,
what measure do you find of
agree or disagree with each of
the following statements? - Who
they have good contacts they manage to get out
ahead."
dat04$pref_effort <- set_labels(dat04$pref_effort,  labels = c("strongly disagree" = 1, "disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "totally agree" = 5))
label(dat04$pref_effort) <- "Thinking of Chilean society,
what measure do you find of
agree or disagree with each of
the following statements? - Who
the more they try they should get
greater rewards than those who
they try less."
dat04$pref_talent <- set_labels(dat04$pref_talent,  labels = c("strongly disagree" = 1, "disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "totally agree" = 5))
label(dat04$pref_talent) <- "Thinking of Chilean society,
what measure do you find of
agree or disagree with each of
the following statements? - Who
possess more talent they should get
greater rewards than those who possess
less talent."
dat04$pref_wpart <- set_labels(dat04$pref_wpart,  labels = c("strongly disagree" = 1, "disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "totally agree" = 5))
label(dat04$pref_wpart) <- "Thinking of Chilean society,
what measure do you find of
agree or disagree with each of
the following statements? - It's fine that those who have rich parents get ahead" 
dat04$pref_netw <- set_labels(dat04$pref_netw,  labels = c("strongly disagree", "disagreement", "Neither agree nor disagree","Agree", "totally agree"))
label(dat04$pref_netw) <- "Thinking of Chilean society,
what measure do you find of
agree or disagree with each of
the following statements? - Is
well that those who have good
contacts get ahead."
```

>>>>

## Translation of variable labels and value labels.
```{r}
label(dat04$grupo) <- "This variable classifies the interviewees according to the order in which the indicators are shown within the scale. There are three orders  of items on the scale. In the first order (group 1), questions of the same factor are then presented to them, in the second order (group 2) they were separated and in the third order (group 3) at random."
dat04$perc_effort <- set_labels(dat04$perc_effort,  labels = c("Strongly disagree" = 1, "Disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "Totally agree" = 5))
label(dat04$perc_effort) <- "Thinking of Chilean society,
what measure do you find of
agree or disagree with each of
the following statements? - Who
the more they try they manage to get bigger
rewards that those who strive
less."
dat04$perc_talent <- set_labels(dat04$perc_talent,  labels = c("Strongly disagree" = 1, "Disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "Totally agree" = 5))
label(dat04$perc_talent) <- "Thinking of Chilean society,
what measure do you find of
agree or disagree with each of
the following statements? - Who
possess more talent they manage to obtain
greater rewards than those who possess
less talent."
dat04$perc_wpart <- set_labels(dat04$perc_wpart,  labels = c("Strongly disagree" = 1, "Disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "Totally agree" = 5))
label(dat04$perc_wpart) <- "Thinking of Chilean society,
what measure do you find of
agree or disagree with each of
the following statements? - Who
they have rich parents manage to get out
ahead."
dat04$perc_netw <- set_labels(dat04$perc_netw,  labels = c("Strongly disagree" = 1, "Disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "Totally agree" = 5))
label(dat04$perc_netw) <- "Thinking of Chilean society,
what measure do you find of
agree or disagree with each of
the following statements? - Who
they have good contacts they manage to get out
ahead."
dat04$pref_effort <- set_labels(dat04$pref_effort,  labels = c("Strongly disagree" = 1, "Disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "Totally agree" = 5))
label(dat04$pref_effort) <- "Thinking of Chilean society,
what measure do you find of
agree or disagree with each of
the following statements? - Who
the more they try they should get
greater rewards than those who
they try less."
dat04$pref_talent <- set_labels(dat04$pref_talent,  labels = c("Strongly disagree" = 1, "Disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "Totally agree" = 5))
label(dat04$pref_talent) <- "Thinking of Chilean society,
what measure do you find of
agree or disagree with each of
the following statements? - Who
possess more talent they should get
greater rewards than those who possess
less talent."
dat04$pref_wpart <- set_labels(dat04$pref_wpart,  labels = c("Strongly disagree" = 1, "Disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "Totally agree" = 5))
label(dat04$pref_wpart) <- "Thinking of Chilean society,
what measure do you find of
agree or disagree with each of
the following statements? - It's fine that those who have rich parents get ahead" 
dat04$pref_netw <- set_labels(dat04$pref_netw,  labels = c("Strongly disagree" = 1, "Disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "Totally agree" = 5))
label(dat04$pref_netw) <- "Thinking of Chilean society,
what measure do you find of
agree or disagree with each of
the following statements? - Is
well that those who have good
contacts get ahead."
```
